---
title: "STAT 601 Final"
Author: Brian Canyon
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

### Introduction

The purpose of this document is to identify any potential changes in typing behavior over time. To try to understand this, we can look at a study performed by Dr. Roy Maxian. He and his colleagues pooled 51 subjects to type a password 50 times a day for 8 consecutive days. The password was ".tie5Roan1". 

### Approach

My approach boils down to looking at change in total typing speed over time. Since their is 400 trails for each user, I created a new variable 'Total Time', which was the mean of all users per trial. My initial impression was to create a linear model that modeled change in overall speed over time. However, I quickly turned away from this idea as a quick look suggested it was not a great fit for this data. The primary reason for this was twofold. Firstly, trials 1-100 saw a significantly greater decrease in time than trials 101-400. Second, it was clear that there were diminishing returns as the user reached around 300 trials. 

These findings resulted in me to turn to polynomial regression. Doing this resulted in a curved line that fit the data much more accurately. That graph can be seen below
```{r echo=FALSE}
### loading in the data
type_data <- read.csv("C:\\Users\\j\\Downloads\\DSL-StrongPasswordData.csv", header = TRUE)

### creating a new vector that represents total time per trial
total_time <- c()
for (x in 1:20400){
  y <- sum(type_data[x,5:35])
  total_time <- c(total_time, y)
}
total_time_cut <- total_time[1:20400]
complete_type_data <- cbind(total_time_cut, type_data)

x <- c(1:400)



average_time_per_trial <- c()
for (x in 1:400){
  average_first_trial <-        mean(complete_type_data$total_time_cut[trial_number == x])
  average_time_per_trial <- c(average_time_per_trial,average_first_trial)
}

a <- 1:400
b <- average_time_per_trial
fit3 <- lm(b ~ a)
fit4 <- lm(b ~ poly(a,4,raw = TRUE))

summary(fit4)

fit4 <- lm(b ~ poly(a,4,raw = TRUE))
plot(1:400, average_time_per_trial, type = "p", col = "green", main = "Typing Speed Change Over Time", ylab = "Total Time Per Trial (Seconds)", xlab = "Repitition Number")
lines(1:400, predict(fit4), col = "blue")

plot(lm(b ~ poly(a,4,raw = TRUE)))
```
### Addition Thoughts

I also wanted to look at the amount of variables (if any) that were not correlated to total speed in any way. The conclusions I found on that front were interesting. Each variable largely fell into two groups. Those groups being either somewhat correlated (cc of .4 to .8) and no correlation (cc of -.2 to .2). 
```{r echo=FALSE}



hist(var_vec, xlab = "Correlation Coefficient", main = "Correlation Coefficient Histogram")
```

### Conclusion

My conclusion comes in two parts. First and most importantly, amount of time shaved off by increased repetitions of typing the same password sees diminishing returns rather quickly. The first 100 trials result in drastic increase in speed while trials 300-400 find very little change in overall speed. It is also interesting to note that not all aspects of typing increase over time. A correlation coefficient around 0 suggests that there is no change in speed over the course of trials.
```{r echo=FALSE}
### First attempt at graphing data to get a look at what is going on with the first user across total time

plot(complete_type_data[1:50,4], complete_type_data[1:50,1], type = 'l', col = 'blue', xlab = "Trail Number", ylab = "Trail Speed", ylim = c(0,11))


### subsetting the larger dataset to isolate a single subject to begin to look for trends and develop a strategy to attack the larger problem

library(tidyverse)
subject_s002 <- complete_type_data %>% filter(ï..subject == 's002')
subject_s003 <- complete_type_data %>% filter(ï..subject == 's003')
subject_s004 <- complete_type_data %>% filter(ï..subject == 's004')
subject_s005 <- complete_type_data %>% filter(ï..subject == 's005')
subject_s006 <- complete_type_data %>% filter(ï..subject == 's006')
subject_s007 <- complete_type_data %>% filter(ï..subject == 's007')


### Below is a graph of average speed over 400 trails for the first subject with both a linear and parabolic regresssion line included



a <- 1:400
b <- average_time_per_trial
fit3 <- lm(b ~ a)
fit4 <- lm(b ~ poly(a,4,raw = TRUE))

summary(fit4)






### Looking at which variables are best correlated to total time.. ie correlation coeficcent of each variable to total time. Creating a for loop to generate a vector of variance between each variable and total time.







```


